Realizing temporal graphs from fastest travel times

### 2023-04-26 New Paper/Realizing TG - extended.tex
MFCS

\begin{definition}[Travel delays]
	Let $(G, \lambda)$ be a $\Delta$-periodic temporal graph.
	Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
	We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
	as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
	More specifically:
	\begin{equation}\label{eq:def-VertexWaitingTime}
		\tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
	\end{equation}
	Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
\end{definition}
Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

@@

\section{Upper bounds in the matrix D}
We start by investigating the hardness of the problem on paths.
We first reduce the input matrix $D$, by performing a series of modifications.

Denote with $P = (v_1, v_2, \dots, v_n)$ a path on $n$ vertices,
and let $D$ be the duration matrix with upper bounds.
For any two vertices $v_i, v_j$ from $P$ we get that:
\begin{itemize}
	\item if the value of $D_{i,j} < D_{i,k}$, where $k \in [1+1,j-1]$
	we set $D_{i,k} = D_{i,j}$.
\end{itemize}
%
By the definition of temporal paths, it follows that $d(i,j) = d(i,k) + d(k-1, j) - 1$, where $i,j \in [n]$, with $i<j$ and $k \in [i+1,j-1]$.
Therefore,
\begin{itemize}
	\item if $D_{i,j} > D_{i,k} + D_{k-1,j} -1$ we set the value of $D_{i,j}$ to $D_{i,k} + D_{k-1,j} -1$. 
\end{itemize}
%
And lastly, to any $i$-th row of the matrix $D$, we make the following modifications:
\begin{itemize}
	\item if there are two entries $j< k$ with the same values, \ie $D_{i,j} = D_{i,k}$ where both $j,k$ are bigger than $i$ (resp.~smaller than $i$),
	then we set $D_{i,j} = D_{i,k} - (k-j)$ (resp.~ $D_{i,k} = D_{i,j} - (k-j)$).
\end{itemize}
We iteratively repeat all of the above steps and produce a reduce matrix $D'$.
It is easy to see that if there exists a labeling of $P$ for $D'$, then this labeling satisfies also $D$.

@@

>>>>

\begin{definition}[Travel delays] \label{def:travelDelays}
	Let $(G, \lambda)$ be a $\Delta$-periodic temporal graph.
	Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
	We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
	as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
	More specifically:
	\begin{equation}\label{eq:def-VertexWaitingTime}
		\tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
	\end{equation}
	Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
\end{definition}
Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

@@

\section{Upper bounds in the matrix D}
We start by investigating the complexity of the problem \DeltaUpperBoundLong\ in the case when the underlying graph $G$, of the input instance, is a path.
To do this efficiently we first present a procedure that reduces the values in the matrix $D$, of the input instance, as much as possible, without effecting solutions.
This procedure is based on the basic properties of fastest temporal paths.

Let $(G,D)$ be an instance of \deltaUpperBound\, where $G$ is a path of length $n$.
Denote with $v_0,v_1, v_2, \dots, v_n$ the vertices of $G$ where $v_{i}v_{i+1} \in E(G)$, for all $i \in [n]$.
Observe the following properties.
\begin{enumerate}
	\item Let $i,j,k \in [n]$, where $i < k < j$, and let $D_{i,j} \leq D_{i,k}$. 
 Then the temporal path $P_{i,k}$ from $v_i$ to $v_k$ cannot be of duration more than $D_{i,j} - (k-j)$,
 which implies that $D_{i,k} \leq D_{i,j} - (k-j)$.

 This holds because $P_{i,k}$ is a subpath of the temporal path $P_{i,j}$ from $v_i$ to $v_j$,
 therefore we can brake down $P_{i,j}$ into two parts, $P_{i,k}$ and the path $P_{k,j}$ from $v_k$ to $v_j$.
 Since $P_{k,j}$ consists of $k-j$ edges we need at least $k-j$ time to traverse it, so the duration of $P_{i,k}$ cannot take more than $D_{i,j} - (k-j)$ time.

    \item From the definition of temporal paths, it follows that $d(i,j) = d(i,k) + d(k-1, j) - 1$, where $i,j \in [n]$, with $i<j$ and $k \in [i+1,j-1]$.
    Therefore $D_{i,j} \leq D_{i,k} + D_{k-1,j} -1$.

    \item 
\end{enumerate}

We iteratively repeat the steps below and produce a reduce matrix $D'$ from the input matrix $D$.
It is easy to see that if there exists a labeling of $P$ for $D'$, then this labeling satisfies also $D$.

@@

\begin{proof}[Proof idea -- new:]
We are given an instance  $(G,\Delta), D$ of \deltaUpperBound\, where $G$ is a path on vertices $v_0, v_1, \dots, v_n$.
Let us denote with $\tau_i^+$ the waiting time at vertex $v_i$ when passing from vertex $v_{i-1}$ to $v_{i+1}$, and with $\tau_i^-$ the waiting time at $v_i$ when traversing through $v_i$ from other direction (for details see~\cref{def:travelDelays}).
Since $\tau_i^- = \Delta - \tau_i^+$, we denote with $\tau_i = \tau_i^+$, and replace $\tau_i^-$ with $ \Delta - \tau_i$.
By the definition it also follows that the duration of a fastest path among two vertices $v_i, v_j$ in $(G,\lambda)$, for $i < j$
is 
\begin{equation}\label{eq:path-positive}
    d(v_i,v_j)=\sum_{k = i+1}^{j-1} \tau_k + 1,
\end{equation}
and for $i > j$ is 
\begin{equation}\label{eq:path-negative}
d(v_i,v_j)=\sum_{k = i+1}^{j-1} (\Delta - \tau_k) + 1.
\end{equation}

We want to determine the values of all $\tau_i$, and consequently the labeling $\lambda$, such that the distance matrix $D$ is satisfied,
namely $d(v_i,v_j) \leq D_{i,j}$ for all $v_i,v_j$ in $G$.
To do this we build an ILP. We do this as follows.

For every pair of vertices $v_i, v_j \in G$ where $|i-j| \geq 2$, we create an inequality of form $d(v_i,v_j) \leq D_{i.j}$, where $d(v_i,v_j)$ equals the expression from \cref{eq:path-positive} if $i < j$, and \cref{eq:path-negative} if $i > j$.
We build a system of $O(n^2)$ inequalities with $n$ variables $d_1,d_2, \dots, d_n$,
where each $d_i \in [\Delta]$. Let us denote this system of inequalities as $I_D$.
We need to prove first that produced system $I_D$ is in fact an (integer) linear program and then show how we can find a solution of it in polynomial time.

Before starting with the analysis of $I_D$ let us observe some basic properties that have to hold for $D$, which will make our computation easier.
\begin{enumerate}
    \item f
\end{enumerate}

\end{proof}

### 2023-04-26 arxivNEW.tex
MFCS

%Due to space constraints, proofs of results marked with $\star$ are (partially) deferred to Appendix.

\subparagraph{Related work.} Graph realization problems on static graphs have been studied since the 1960s. We provide an overview of the literature in the introduction.

To the best of our knowledge, we are the first to consider graph realization problems in the temporal setting. However, many other connectivity-related problems have been studied in the temporal setting~\cite{Mertzios-transitivity21,Akrida-explorer-21,enright2021deleting,MolterRZ21,klobas2023interference,deligkas2022optimizing,erlebach2021temporal,Flu+19a,Zsc+19,CasteigtsCS22,FuchsleMNR22}, most of which are much more complex and computationally harder than their non-temporal counterparts, and some of which do not even have a non-temporal counterpart.

There are two problem settings that share similarities with our setting, which we discuss in more detail in the following.

Several problems have been studied where the goal is to assign labels to (sets of) edges of a given static graph in order to achieve certain connectivity-related properties~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity,enright2021assigning}. The main difference to our problem setting is that in the mentioned works, the input is a graph and the sought labeling is not periodic. Furthermore, the investigated properties are temporal connectivity among all vertices~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity}, temporal connectivity among a subset of vertices~\cite{KlobasMMS22}, or reducing reachability among the vertices~\cite{enright2021assigning}. In all cases the duration of the temporal connections is not of interest.

The only work we are aware of that studies problems on periodic temporal graphs is work on the temporal cops and robber game~\cite{Morawietz021,MorawietzRW20,ErlebachS20}. The question here is (in its basic case) whether a certain number of cops are able to catch a robber, if everybody can move along one edge at every time step and the robber is considered caught if it occupies the same vertex as a cop at the same time.

\subparagraph{Preliminaries and notation.}
We already introduced the most central notion and concepts. There are some additional definitions we need to present our proofs and results which we give in the following.

@@

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$. 
For simplicity of the presentation~--and with a slight abuse of notation--~we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as 
$D_{a,b}$, where $a=v_i$ and $b=v_j$. That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ whenever it is clear from the context.
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda,\Delta)$,
where we actually mean that $(P,\lambda,\Delta)$ is a temporal path with $P$ as its underlying (static) path.
We denote with $d(u,v,P)$ the duration of the fastest temporal path from $u$ to $v$ along path $P$. Note that $d(u,v,P) \geq d(u,v)$.

@@

\begin{comment}
Let $(G,\lambda)$ be a temporal graph with  $n = |V(G)|$ and  $m=|E(G)|$.
Let $v \in V(G)$ be an arbitrary vertex of $G$, denote with $deg_{max}$ the maximum degree of all vertices in $V(G)$ and let $S_v$ be the set of time edges in $(G,\lambda)$, incident to $v$, \ie $S_v = \{(uv, \lambda(uv)) | uv \in E(G)\}$.
In their work Wu et~al.~\cite{Wu2016Efficient} provide an algorithm that computes the duration of the fastest temporal path from a source vertex $v$ to all other vertices in $G$ in
$O(n + m \log c)$ time, where $c = \min \{|S_v|, deg_{max}\}$.
From this result, we can deduce the following. 
\begin{lemma} \label{lemma:calculatingD}
Let $(G, \lambda, \Delta)$ be a $\Delta$-periodic temporal graph, where $\lambda$ assigns exactly $k$ labels to every edge of $G$. 
Denote with $deg_{max}$ the maximum degree of vertices in $G$, let $deg(v_i)$ be the degree of vertex $v_i \in V(G)$,
and let $c_{v_i} = \min \{ k \cdot deg(v_i), deg_{max}\}$.\\
To calculate the durations of all fastest temporal paths from a fixed vertex $v \in V(G)$ to all other vertices, it takes $O(n + m \log c_v)$ time. \\
To calculate the durations of the fastest temporal paths among all pairs of vertices we then need $O(n^2 + m \sum_{v \in V(G)} \log c_v)$ time.
\end{lemma}

Observe that $O(n + m \log c_v) = O(n + m \log deg_{max}) = O(n + m \log n) = O(n^2 \log n)$ and similarly
$O(n^2 + m \sum_{v \in V(G)} \log c_v) = O(n^2 + mn \log deg_{max}) = O(n + mn \log n) = O(n^3 \log n)$.
Since an instance $I$ of \kDeltaExact\ problem is an $n \times n$ matrix $D$, the running time of calculating one row of $D$ (\ie the fastest temporal paths from a single vertex to all others) takes $O(|I|\log \sqrt{|I|})$-time and calculating the whole matrix $D$ (\ie the fastest temporal paths among all pairs of vertices) takes $O(|I|^{3/2} \log \sqrt{|I|})$ time.
Therefore, from now on we say that calculating (a row of) $D$ takes polynomial time.
\end{comment}
%\section{Results}

@@

%
\begin{figure}[t]
	\noindent
	\makebox[\textwidth]{
		\centering
	\includegraphics{fig-NP-reduction}
}
\caption{Illustration of the temporal graph $(G,\lambda)$ from the NP-hardness reduction, 
	where the NAE 3-SAT formula $\phi$ is of form $\phi = \text{NAE}(x_1, \overline{x}_2, x_3) \wedge \text{NAE}(\overline{x}_1, x_2, x_4)$.
To improve the readability, we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with gray dotted lines.
Presented is the labeling of $G$ corresponding to the assignment $x_1=x_2=\textsc{true}$ and $x_3,x_4=\textsc{false}$,
where all unlabeled edges get the label $2$.
}\label{fig:NP-example}
\end{figure}
%
We set $\Delta$ to some constant larger than two, that is, $\Delta\ge 3$. Next, we specify the durations in the matrix $D$ between all vertex pairs.
For the sake of simplicity we write $D_{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following durations:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following durations:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following duratios:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following duratios:
    $d(x_i,c)=\Delta$, $d(c,x_i)=2\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following duratios:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following duratios:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of the instance $(D,\Delta)$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $(D,\Delta)$ is a \textsc{Yes}-instance of \deltaExact\ if and only if NAE 3-SAT formula $\phi$ is satisfiable.

@@

    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we have to check that
    first,
    $d(c,x_i^T)=d(c,x_i^F)=2$,
    second, $d(x_i^T, c) = d(x_i^F,c)=\Delta$,
    third, $d(x_i,c)=\Delta$,
    and fourth $d(c,x_i)=2 \Delta + 2$.

    Let $x_j$ be a variable that appears non-negated in $c$ (the case where $x_j$ appears negated is symmetric). 
    Then there is a temporal path 
    from $c$ to $x_i^T$ via $x_j^T$ and also a temporal path 
    from $c$ to $x_i^F$ via $x_j^T$ such that $\lambda(x_j^T c)=1$ and $\lambda(x_j^T x_i^T)=\lambda(x_j^T x_i^F)=2$,
    which proves the first equality.
    Using the same temporal path in the opposite direction,
    \ie first the edge $x_j^T c$ and then one of the edges $x_j^T x_i^F$ or $x_j^T x_i^T$,
    yields the second equality.
    For a temporal path from $c$ to $x_i$ we traverse the following three edges 
    $x_j^T c, x_j^T x_i^F$ and $x_i^Fx_i$,
    with labels $1, 2, 2$ respectively.
    This proves the second equality.
    Now for the case of a temporal path from $x_i$ to $c$,
    we use the same three edges, but in the opposite direction,
    namely $x_i^Fx_i, x_j^T x_i^F$ and $x_j^T c$,
    again with labels $2, 2, 1$, respectively,
    which proves the last equality.
    Note that all of the above temporal paths are also the shortest possible, 
    and since the labels of first and last edges (of these paths) are unique,
    it follows that we cannot find faster temporal paths.

@@

    Let $x_k$ be a variable that appears non-negated in $c_i$
    and $x_\ell$ the variable that appears non-negated in $c_j$
    (all other cases are symmetric).
    There is a path of length three from $c_i$ to $c_j$
    that passes first through vertex $x_k^T$ and then through vertex $x_\ell^T$.
    Therefore the temporal path from $c_i$ to $c_j$
    uses the edges $x_k^T c_i, x_\ell ^T c_j, x_k^T x_\ell ^T$, with labels $1,2,1$, respectively,
    which proves the desired equality.
    Note also that this is the shortest path between $c_i$ and $c_j$, and that the first and the last edge must have the label $1$,
    therefore it follows that this is the fastest temporal path.
\end{itemize}
Lastly, observe that the above constructed labeling $\lambda$ uses values $\{1,2,3\}\subseteq[\Delta]$, since we have $\Delta\ge 3$.
\end{proof}

@@

\begin{figure}%[t]
\noindent\makebox[\textwidth]{
\centering
\includegraphics{fig-W1-reduction}
    }
    \caption{Illustration of part of the underlying graph $G$ and a possible labeling. Edges incident with vertices of connector gadgets are omitted. Gray vertices form a feedback vertex set.
    The double line connections, between a vertex $v_{i-1}^j$ in the verification gadget, and $u_1^3$ in the edge selection gadget, 
    and, between a vertex $u_2^3$ in the edge selection gadget, and $v_{i}^j$ in the verification gadget,
    consist of $5n$ vertices $a_1^{j,i,3},a_2^{j,i,3},\dots,a_{5n}^{j,i,3}$ 
    and $b_1^{j,i,3},b_2^{j,i,3},\dots,b_{5n}^{j,i,3}$, respectively.
    }\label{fig:hardness1}
\end{figure}

@@

    From the above definition and the definition of the duration of the temporal path $P$ we get the following two observations.
    \begin{observation}\label{obs:durationPwithWaitingTimes}
        Let $P = (v_0, v_1, \dots, v_p)$ be the underlying path of the temporal path
        $(P, \lambda)$ from $v_0$ to $v_p$.
        Then $d(P,\lambda) = \sum_{i = 1}^{p-1} \tau_{v_i}^{v_{i-1}v_i} + 1 $.
    \end{observation}
    \begin{proof}
        For the simplicity of the proof denote $t_i = \lambda(v_{i-1}v_i)$, and suppose that $t_i \leq t_{i+1}$, for all $i \in \{1,2,3,\dots,p\}$.
        Then
        \begin{align*}
        \sum_{i = 1}^{p-1} \tau_{v_i}^{v_{i-1}v_i} + 1  
        &= \sum _{i = 1}^{p-1} (t_{i+1} - t_i) + 1 \\
        & = (t_2 - t_1) + (t_3 - t_2) + \cdots + (t_p - t_{p-1}) + 1  \\
        & = t_{p-1} - t_1 + 1\\
        & = d(P)
        \end{align*}
        Now in the case when $t_i > t_{i+1}$ we get that $\tau_{v_i}^{v_{i-1}v_{i+1}} = \Delta + t_{i+1} - t_i$.
        At the end this still results in the correct duration as the last time we traverse the path $P$ is not exactly $t_p$ but $k \lambda + t_p$, for some $k$.
    \end{proof}
    We also get the following.
    \begin{observation}\label{obs:travel-delays-both-directions}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of problem \deltaExact.
    For any two incident edges $e_1 = uv$ and $e_2 = vz$ on vertices $u,v,z \in V$, with $e_1 \cap e_2 = v$, we have $\tau_v^{zu} = \Delta - \tau_v^{uz} \pmod \Delta$.
    \end{observation}

@@

\begin{lemma} \label{lemma:unique-fastest-path-allDelays}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of problem \deltaExact,
    and let $(P,\lambda)=(P_{1,k},\lambda)$ be a fastest temporal path from $u=v_1$ to $v=v_k$ on $k$ vertices $v_1,v_2,\dots,v_k$.
    Let us denote with $(P_{1,i}, \lambda)$ the sub-path of temporal path $(P_{1,k},\lambda)$, that starts at $v_1$ and finishes at $v_i$.
    Suppose that for any $v_i$ in $P$ it follows that $(P_{1,i}, \lambda)$ is also the fastest temporal path from $u=v_1$ to $v_i$.
    Then we can determine travel delays on vertices of $P$ using the following equation
    \begin{equation}\label{eq:trav-delays-paths}
        \tau_{v_i}^{v_{i-1},v_{i+1}} = D_{1,i+1} - D_{1,i},
    \end{equation}
    for all $i \in \{2,3, \dots, k-1\}$.
\end{lemma}

>>>>

Due to space constraints, proofs of results marked with $\star$ are (partially) deferred to the Appendix.

\subparagraph{Related work.} Graph realization problems on static graphs have been studied since the 1960s. We provide an overview of the literature in the introduction. 
To the best of our knowledge, we are the first to consider graph realization problems in the temporal setting. However, many other connectivity-related problems have been studied in the temporal setting~\cite{Mertzios-transitivity21,Akrida-explorer-21,enright2021deleting,MolterRZ21,klobas2023interference,deligkas2022optimizing,erlebach2021temporal,Flu+19a,Zsc+19,CasteigtsCS22,FuchsleMNR22}, most of which are much more complex and computationally harder than their non-temporal counterparts, and some of which do not even have a non-temporal counterpart.

There are some problem settings that share similarities with ours, which we discuss now in more detail.

Several problems have been studied where the goal is to assign labels to (sets of) edges of a given static graph in order to achieve certain connectivity-related properties~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity,enright2021assigning}. The main difference to our problem setting is that in the mentioned works, the input is a graph and the sought labeling is not periodic. Furthermore, the investigated properties are temporal connectivity among all vertices~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity}, temporal connectivity among a subset of vertices~\cite{KlobasMMS22}, or reducing reachability among the vertices~\cite{enright2021assigning}. In all these cases, the duration of the temporal paths has not been considered.

%work on the temporal cops and robber game~\cite{morawietz2021timecop,morawietz2020timecop,ErlebachS20}. The question here is (in its basic case) whether a certain number of cops are able to catch a robber, if everybody can move along one edge at every time step and the robber is considered caught if it occupies the same vertex as a cop at the same time. Several other works on periodic temporal graphs were mentioned in the introduction.

Finally, there are many models for dynamic networks in the context of distributed computing~\cite{Kuhn2011Dynamic}. 
These models have some similarity to temporal graphs, in the sense that in both cases the edges appear and disappear over time. 
However, there are notable differences. For example, one important assumption in the distributed setting 
can be that the edge changes are adversarial or random (while obeying some constraints such as connectivity), 
and therefore they are not necessarily known in advance~\cite{Kuhn2011Dynamic}.

\subparagraph{Preliminaries and notation.}
We already introduced the most central notion and concepts. There are some additional definitions we need, to present our proofs and results which we give in the following.

@@

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$. 
For simplicity of the presentation~--and with a slight abuse of notation--~we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as 
$D_{a,b}$, where $a=v_i$ and $b=v_j$. That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ whenever it is clear from the context.
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda,\Delta)$,
where we actually mean that $(P,\lambda,\Delta)$ is a temporal path with $P$ as its underlying (static) path.

@@

@@

%
\begin{figure}[t]
	\noindent
	\makebox[\textwidth]{
		\centering
	\includegraphics{fig-NP-reduction}
}
\caption{Illustration of the temporal graph $(G,\lambda)$ from the NP-hardness reduction, 
	where the NAE 3-SAT formula $\phi$ is of form $\phi = \text{NAE}(x_1, \overline{x}_2, x_3) \wedge \text{NAE}(x_1, x_2, x_4)$.
To improve the readability, we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with gray dotted lines.
Presented is the labeling of $G$ corresponding to the assignment $x_1=x_2=\textsc{true}$ and $x_3,x_4=\textsc{false}$,
where all unlabeled edges get the label $2$.
}\label{fig:NP-example}
\end{figure}
%
We set $\Delta$ to some constant larger than two, that is, $\Delta\ge 3$. Next, we specify the durations in the matrix $D$ between all vertex pairs.
For the sake of simplicity we write $D_{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following durations:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following durations:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following duratios:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following duratios:
    $d(x_i,c)=2 \Delta$, $d(c,x_i)=\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following duratios:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following duratios:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of the instance $(D,\Delta)$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $(D,\Delta)$ is a \textsc{Yes}-instance of \deltaExact\ if and only if NAE 3-SAT formula $\phi$ is satisfiable.

@@

    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we have to check that
    first,
    $d(c,x_i^T)=d(c,x_i^F)=2$,
    second, $d(x_i^T, c) = d(x_i^F,c)=\Delta$,
    third, $d(c,x_i)=\Delta + 2$,
    and fourth $d(x_i,c)=2 \Delta$.

    Let $x_j$ be a variable that appears non-negated in $c$ (the case where $x_j$ appears negated is symmetric). 
    Then there is a temporal path 
    from $c$ to $x_i^T$ via $x_j^T$ and also a temporal path 
    from $c$ to $x_i^F$ via $x_j^T$ such that $\lambda(x_j^T c)=1$ and $\lambda(x_j^T x_i^T)=\lambda(x_j^T x_i^F)=2$,
    which proves the first equality.
    Using the same temporal path in the opposite direction,
    \ie first the edge $x_j^T c$ and then one of the edges $x_j^T x_i^F$ or $x_j^T x_i^T$ at times $2$ and $\Delta + 1$, respectively,
    yields the second equality.
    For a temporal path from $c$ to $x_i$ we traverse the following three edges 
    $x_j^T c, x_j^T x_i^F$ and $x_i^Fx_i$,
    with labels $1, 2, 2$ respectively (\ie the path traverses them at time $1,2$ and $\Delta + 2$, respectively), which proves the third equality.
    Now for the case of a temporal path from $x_i$ to $c$,
    we use the same three edges, but in the opposite direction,
    namely $x_i^Fx_i, x_j^T x_i^F$ and $x_j^T c$,
    again at times $2, \Delta + 2, 2\Delta + 1$, respectively,
    which proves the last equality.
    Note that all of the above temporal paths are also the shortest possible, 
    and since the labels of first and last edges (of these paths) are unique,
    it follows that we cannot find faster temporal paths.

@@

    Let $x_k$ be a variable that appears non-negated in $c_i$
    and $x_\ell$ the variable that appears non-negated in $c_j$
    (all other cases are symmetric).
    There is a path of length three from $c_i$ to $c_j$
    that passes first through vertex $x_k^T$ and then through vertex $x_\ell^T$.
    Therefore the temporal path from $c_i$ to $c_j$
    uses the edges $x_k^T c_i, x_\ell ^T c_j, x_k^T x_\ell ^T$, with labels $1,2,1$ (at times $1, 2, \delta + 1$), respectively,
    which proves the desired equality.
    Note also that this is the shortest path between $c_i$ and $c_j$, and that the first and the last edge must have the label $1$,
    therefore it follows that this is the fastest temporal path.
\end{itemize}
Lastly, observe that the above constructed labeling $\lambda$ uses values $\{1,2,3\}\subseteq[\Delta]$, therefore $\Delta\ge 3$.
\end{proof}

@@

\begin{figure}%[t]
\noindent\makebox[\textwidth]{
\centering
\includegraphics{fig-W1-reduction}
    }
    \caption{Illustration of part of the underlying graph $G$ and a possible labeling. Edges incident with vertices $\hat{v}_1,\hat{v}_2$ of connector gadgets are omitted. Gray vertices form a feedback vertex set.
    The double line connections, between a vertex $v_{i-1}^j$ in the verification gadget, and $u_1^3$ in the edge selection gadget, 
    and, between a vertex $u_2^3$ in the edge selection gadget, and $v_{i}^j$ in the verification gadget,
    consist of $5n$ vertices $a_1^{j,i,3},a_2^{j,i,3},\dots,a_{5n}^{j,i,3}$ 
    and $b_1^{j,i,3},b_2^{j,i,3},\dots,b_{5n}^{j,i,3}$, respectively.
    }\label{fig:hardness1}
\end{figure}

@@

    From the above definition and the definition of the duration of the temporal path $P$ we get the following two observations.
    \begin{observation}\label{obs:durationPwithWaitingTimes}
        Let $P = (v_0, v_1, \dots, v_p)$ be the underlying path of the temporal path
        $(P, \lambda)$ from $v_0$ to $v_p$.
        Then $d(P,\lambda) = \sum_{i = 1}^{p-1} \tau_{v_i}^{v_{i-1}v_i} + 1 $.
    \end{observation}
    \begin{proof}
        For the simplicity of the proof denote $t_i = \lambda(v_{i-1}v_i)$, and suppose that $t_i \leq t_{i+1}$, for all $i \in \{1,2,3,\dots,p\}$.
        Then
        \begin{align*}
        \sum_{i = 1}^{p-1} \tau_{v_i}^{v_{i-1}v_i} + 1  
        &= \sum _{i = 1}^{p-1} (t_{i+1} - t_i) + 1 \\
        & = (t_2 - t_1) + (t_3 - t_2) + \cdots + (t_p - t_{p-1}) + 1  \\
        & = t_{p-1} - t_1 + 1\\
        & = d(P, \lambda)
        \end{align*}
        Now in the case when $t_i > t_{i+1}$ we get that $\tau_{v_i}^{v_{i-1}v_{i+1}} = \Delta + t_{i+1} - t_i$.
        At the end this still results in the correct duration as the last time we traverse the path $P$ is not exactly $t_p$ but $k \lambda + t_p$, for some $k$.
    \end{proof}
    We also get the following.
    \begin{observation}\label{obs:travel-delays-both-directions}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of the \deltaExact problem.
    For any two incident edges $e_1 = uv$ and $e_2 = vz$ on vertices $u,v,z \in V$, with $e_1 \cap e_2 = v$, we have $\tau_v^{zu} = \Delta - \tau_v^{uz} \pmod \Delta$.
    \end{observation}

@@

\begin{lemma} \label{lemma:unique-fastest-path-allDelays}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of the \deltaExact problem,
    and let $(P,\lambda)=(P_{1,k},\lambda)$ be a fastest temporal path from $u=v_1$ to $v=v_k$ on $k$ vertices $v_1,v_2,\dots,v_k$.
    Let us denote with $(P_{1,i}, \lambda)$ the sub-path of temporal path $(P_{1,k},\lambda)$, that starts at $v_1$ and finishes at $v_i$.
    Suppose that for any $v_i$ in $P$ it follows that $(P_{1,i}, \lambda)$ is also the fastest temporal path from $u=v_1$ to $v_i$.
    Then we can determine travel delays on vertices of $P$ using the following equation
    \begin{equation}\label{eq:trav-delays-paths}
        \tau_{v_i}^{v_{i-1},v_{i+1}} = D_{1,i+1} - D_{1,i},
    \end{equation}
    for all $i \in \{2,3, \dots, k-1\}$.
\end{lemma}
